--- 
title: "NYC 2022 Shooting Incident Analysis"
author: "William Gu, Shujie Hu"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---
--- 
title: "NYC 2022 Shooting Incident Analysis"
author: "William Gu, Shujie Hu"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
---
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

# Introduction

  As shooting incidents are one of the most fatal crimes but also one that happens frequently in NYC, we’d like to investigate the correlation behind these incidents, hoping to explore the patterns of New York City’s shooting incidents to study perpetrator-victim relationships and address personal safety awareness. Since crime statistics may differ from the years prior to the pandemic, we will explore the most recent shooting incidents in NYC.
  Through various data visualization techniques, we hope to answer the following questions that we are interested in studying:
<br>
1. The Number of Shooting Incidents With Respect To Month (Line Chart with LOESS Smoother)<br>
2. The Ratio of Death in the Shooting Incidents With Respect to Hours In a Day. (Stacked Bar Chart) <br>
3. The Relationship Between the Month and Boroughs of Shooting Incidents (Line Chart & Stacked Histogram)<br>
4. The Ratio of Death With Respect To The Specific Location of The Shooting Incidents (Cleveland dot plot)<br>
5. The Ratio of Death For The Shooting Incidents In Each Borough (Stacked Bar Chart)<br>

6. The Number of Shooting Incidents With Respect to the Precincts. (Map) <br>
7. The Correlation Between Different Boroughs and the Victims/Perpetrators. (Histogram + facet)<br>
8. The Contrast Between Perpetrators' Age vs Victims’ Age. (Population Pyramid) <br>
9. The Visualization of the Victims’ Characteristics. (Mirror Chart + facet & Alluvial Chart)<br>
10. The Visualization of the Perpetrators' Characteristics (PERP_AGE_GROUP, PERP_SEX, and PERP_RACE). (Spine Plot)<br>
 
browseURL("docs/index.html")

<!--chapter:end:index.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Proposal

## Research topic

Behind the beautiful skyline and bustling streets, criminal scenes may happen on every corner at any moment. The consecutive murder incidents that happened this summer and the shooting in the East Village right after the Fall semester starts caused panic among the citizens. Public safety has always been a concern, especially for the ones who just moved to New York City. Though these shooting incidents seem to be random at first glance, data analysis and visualization help reveal the hidden correlation and provide hints about public safety concerns. <br><br>

The New York Police Department records reported crimes and releases data sets related to police enforcement and criminal activity, aiming to increase transparency and foster collaboration. With the data provided by NYPD, we are interested in the citywide crime statistics, specifically shooting incidents in 2022. It is worth studying these up-to-date shooting incidents when living in NYC, thereby we can have a clear picture of the crimes around us to avoid possible dangers. Specifically, we are interested in questions relating to the characteristics of victims and suspects, and the possible correlation between the victims/suspects' characteristics with respect to the incident's geography or occurrence time. 
 


## Data availability

  The [datasource](https://www1.nyc.gov/site/nypd/stats/crime-statistics/citywide-crime-stats.page) of this project is from the data published and maintained by the New York Police Department. There were many other data sources that we considered when deciding to analyze crime-related information about New York City, however, the New York Police Department is the most reliable source as their data is collected based on the reported crimes they record. Therefore, we considered collecting the citywide crime data directly released by the New York Police Department. <br><br>
NYPD provides quarterly-updated crime incident datasets such as arrest and complaint, among which we are most interested in and therefore decide to analyze the shooting incidents this year. We may also consider referring to statistics of other crimes or shootings in former years for comparison.<br><br>
  The New York Police Department provides the shooting incidents data in multiple formats on their official website, among these formats, we found csv to be the most convenient to be analyzed in R. The next steps would be importing the data from the NYPD website and performing data pre-processing such as combining datasets and cleaning by using either Python or R.<br><br>
  Overall, the data is reliable and of good quality. However, there are null values for certain columns, for example, LOCATION_DESC and information about the perpetrator like PERP_AGE_GROUP and PERP_SEX. Though the missing information about the perpetrators is understandable as the police may not be able to arrest the perpetrator in the crime scene, we will check these missing values specifically with a missing value analysis. We will pay special consideration regarding missing values in different situations in our project, for some tasks we may need to neglect these missing values as the information is not provided to us, while for some tasks about the perpetrator we may regard these missing values as a separate group and perform analysis on this group.

<!--chapter:end:proposal.Rmd-->


# Data 

Placeholder


## Sources
## Cleaning / transformation
## Missing value analysis

<!--chapter:end:data.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Results

```{r}
#install.packages("vcdExtra")
#install.packages("sf")
#install.packages("tmap")
#install.packages("ggalluvial")
#install.packages("plotly")
```

```{r}
library(tidyverse)
library(ggplot2)
library(lubridate)
```

```{r}
data <- read.csv(file = 'data/NYPD_Shooting_Incident_Data__Year_To_Date_.csv', na.strings = "(null)")
# head(data)
```
## The Victims/Perpetrators' Characteristics

The Contrast Between Perpetrators' Age vs Victims’ Age (Population Pyramid):
```{r}
library(tibble)
# As the perpetrator may escape and results in an empty record, the NAs are removed
perp <- na.omit(data) 

hist_perp <- perp %>%
  group_by(PERP_AGE_GROUP) %>%
  dplyr::summarize(PERP_Freq = n()) %>%
  ungroup() %>%
  complete(PERP_AGE_GROUP, fill = list(PERP_Freq = 0))
  colnames(hist_perp) <- c('AGE_GROUP','Freq')
  hist_perp <- hist_perp %>%
  tibble::add_column(Type = "Perpetrator")

hist_vic <- perp %>%
  group_by(VIC_AGE_GROUP) %>%
  dplyr::summarize(VIC_Freq = n()) %>%
  ungroup() %>%
  complete(VIC_AGE_GROUP, fill = list(VIC_Freq = 0))
  colnames(hist_vic) <- c('AGE_GROUP','Freq')
  hist_vic <- hist_vic %>%
  tibble::add_column(Type = "Victim")

pyramid <- rbind(hist_perp, hist_vic)
pyramid <- pyramid %>% mutate(Freq = ifelse(Type=="Perpetrator", Freq*(-1), Freq*1))
color <- c("darkblue", "darkred")

ggplot(pyramid, aes(x = AGE_GROUP, y = Freq, fill=Type)) +
    geom_bar(stat = "identity") +
    coord_flip()+
    scale_fill_manual(values = color) +
    labs(title = "Population Pyramid", x = "Age Group", y = "Population") +
    theme_grey(13)
```

Note: In order to present the age groups in its order, this pyramid is not sorted based on the population(count)
<br>
From this graph, it is obvious that the age group 25-44 ranks first in the number of both perpetrators and victims. However, it is also worth noticing that the number of teenage perpetrators(age < 24) is more than that of the teenage victims, whereas this difference is reversed for the age groups above 25, so that there are more victims than perpetrators for ages above 25, especially for ages above 45. Thus speaking of the age group distribution, perpetrators are younger than the victims.   


<br><br>
The Visualization of the Victims’ Characteristics (Mirror Chart):
```{r}
vic <- na.omit(data) 

mirror <- vic %>%
  group_by(VIC_AGE_GROUP, VIC_SEX, VIC_RACE) %>%
  dplyr::summarize(Freq = n()) %>%
  ungroup() %>%
  complete(VIC_AGE_GROUP, VIC_SEX, VIC_RACE, fill = list(Freq = 0))

mirror <- mirror %>% mutate(Freq = ifelse(VIC_SEX=="F", Freq*(-1), Freq*1)) 

ggplot(mirror, aes(x = VIC_AGE_GROUP, y = Freq, fill = VIC_SEX)) + 
    geom_bar(stat = "identity") +
    facet_wrap(~VIC_RACE) +
    coord_flip() +
    scale_y_continuous(limits = c(-25, 125), breaks = seq(-25, 125, by = 50)) +
    labs(title = "Mirror Chart of the Victims' Age Group With Respect to Gender in Different Races", x = "Age Group", y = "Number of Victims") +
    theme_grey(12)
```

From the mirror chart above, among all the victims, Black has the highest number of victims in terms of race, while the age group 25-44 ranks top among all races except for White. Most of the victims are male, but there are female victims across almost all age groups and all races except for Asian/Pacific Islander.

<br><br>
Victims’ Characteristics Alternative (Alluvial Chart)
```{r}
library(ggalluvial)

df_allu <- data |> filter(VIC_AGE_GROUP != "1022" ) %>% na.omit()
df_allu <- df_allu |> mutate(VIC_AGE_GROUP = case_when(
  VIC_AGE_GROUP == "45-64" ~ "45+",
  VIC_AGE_GROUP == "65+" ~ "45+",
  TRUE ~ VIC_AGE_GROUP
))
df_allu <- df_allu |> mutate(VIC_RACE = case_when(
  VIC_RACE == "AMERICAN INDIAN/ALASKAN NATIVE" ~ "ASIAN/AMERICAN INDIAN",
  VIC_RACE == "ASIAN / PACIFIC ISLANDER" ~ "ASIAN/AMERICAN INDIAN",
  VIC_RACE == "BLACK HISPANIC" ~ "HISPANIC",
  VIC_RACE == "WHITE HISPANIC" ~ "HISPANIC",
  TRUE ~ VIC_RACE  
))
  
df_allu <- df_allu |> group_by(VIC_AGE_GROUP, VIC_SEX, VIC_RACE) |> summarize(Freq = n())

ggplot(df_allu, aes(y=Freq, axis1=VIC_SEX, axis2=VIC_AGE_GROUP, axis3=VIC_RACE ))+
  geom_alluvium(aes(fill=VIC_AGE_GROUP), width = 1/12) +
  geom_stratum(width = 1/12, fill="grey80", color="grey") +
  geom_label(stat = "stratum",
             aes(label = after_stat(stratum))) +
  scale_x_discrete(expand = c(.05, .05)) +
  scale_fill_brewer(type = "qual", palette = "Set1") +
  ggtitle("Victims Characteristics") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_void()
```
Similar to the previous findings, male victims are a lot more than female victims. Most of them are within 18-24 and 25-44 age groups, which are young adults and mid-age people. Black people are the majority races, followed by Hispanics. White, Asian and other race are the minorities in terms of victims.

<br><br>
The Visualization of the Perpetrators' Characteristics (PERP_AGE_GROUP, PERP_SEX, and PERP_RACE). (Spine Plot)
```{r}
library("plyr")
count(data, 'PERP_RACE')
```

```{r, results='hold'}
library(vcdExtra)
library(forcats)

perp <- data %>% 
  mutate(PERP_AGE_GROUP = fct_relevel(PERP_AGE_GROUP, "<18"),
         PERP_RACE = fct_relevel(PERP_RACE, "WHITE","ASIAN / PACIFIC ISLANDER","BLACK HISPANIC","WHITE HISPANIC","BLACK"),
         PERP_SEX = fct_relevel(PERP_SEX, "M"))

vcd::mosaic(PERP_RACE ~ PERP_SEX + PERP_AGE_GROUP, perp,
       direction = c("v", "v", "h"),
       highlighting_fill= RColorBrewer::brewer.pal(5, "Accent"), 
       labeling_args = list(rep = c(PERP_AGE_GROUP = FALSE), 
       abbreviate = c(PERP_RACE = 1),
       rot_labels = c(30, 30, 30, 30), 
       offset_varnames = c(right = 1), 
       offset_labels = c(right = 0.3)), 
       margins = c(right = 4, bottom = 3)
       )
```


Note: To make the overlapped labels readable, (B,WH,BH,A/PI,W) on the y-axis are the abbreviations for "BLACK", "WHITE HISPANIC", "BLACK HISPANIC", "ASIAN / PACIFIC ISLANDER", "WHITE".
<br>
There are significantly more male perpetrators than female perpetrators, with 25-44 ranks first in terms of age group and Black in terms of race. There is also a correlation with the former graph of the victims in terms of the rank of age group and race.

## Visualization Relating to Incidents Demographics 

The Number of Shooting Incidents With Respect to the Precincts (Map):
```{r}
library(sf)
library(tmap)

ny_police = st_read('data/nyc_precinct/nypp.shp',quiet=TRUE) 

ny_crime <- data %>% 
  group_by(PRECINCT) %>%
  dplyr::summarize(Freq = n()) %>%
  ungroup() %>%
  complete(PRECINCT, fill = list(Freq = 0)) %>%
  dplyr::rename(Precinct = PRECINCT) 

ny_crime <- left_join(ny_police, ny_crime)
ny_crime[is.na(ny_crime)] <- 0

ny_crime %>%
  tm_shape() +
  tm_polygons("Freq", palette = "Reds", title="") +
  tm_text("Precinct", size = .65) +
  tm_layout("Number of Shooting Incidents \nby NYC Police Precinct in 2022",
            title.size = .95, frame = FALSE)
```

Referring to the map of NYC, Precinct 44 is the Bronx, Precinct 75 is East New York, and Precinct 73 is Brownsville, the three areas where most shooting incidents occurred this year.

<br><br>
The Correlation Between Different Boroughs and the Victims/Perpetrators. (Histogram + facet)
```{r}
# vic <- na.omit(data)
hist_vic <- vic %>%
  group_by(BORO, VIC_AGE_GROUP, VIC_RACE) %>%
  dplyr::summarize(Freq = n()) %>%
  ungroup() %>%
  complete(BORO, VIC_AGE_GROUP, VIC_RACE, fill = list(Freq = 0))

ggplot(hist_vic, aes(x = VIC_AGE_GROUP, y = Freq, fill = VIC_RACE)) +
  geom_col(position = "dodge") +
  facet_wrap(~BORO) + 
  scale_y_continuous(limits = c(0, 50), breaks = seq(0, 50, by = 10)) +
  scale_x_discrete(guide = guide_axis(n.dodge=3)) + 
  scale_fill_brewer(palette = "Set2") + 
  labs(title = "The Victims' Characteristics With Respect to the Five Boroughs", x = "Victims' Age Group", y = "Number of Victims") +
  theme_grey(13)

# perp <- na.omit(data) 

hist_perp <- perp %>%
  group_by(BORO, PERP_AGE_GROUP, PERP_RACE) %>%
  dplyr::summarize(Freq = n()) %>%
  ungroup() %>%
  complete(BORO, PERP_AGE_GROUP, PERP_RACE, fill = list(Freq = 0))

ggplot(hist_perp, aes(x = PERP_AGE_GROUP, y = Freq, fill = PERP_RACE)) +
  geom_col(position = "dodge") +
  facet_wrap(~BORO) + 
  scale_y_continuous(limits = c(0, 50), breaks = seq(0, 50, by = 10)) +
  scale_x_discrete(guide = guide_axis(n.dodge=3)) + 
  scale_fill_brewer(palette = "Set2") + 
  labs(title = "The Perpetrators' Characteristics With Respect to the Five Boroughs", x = "Perpetrators' Age Group", y = "Number of Perpetrators") +
  theme_grey(13)

```

In general, Brooklyn and Manhattan are the boroughs where most shooting incidents occurred. Most of the perpetrators and victims are under the age group 25-44 in all boroughs, however, there are more teenage perpetrators (18-24) in some of the boroughs like the Bronx, while the number of victims under the age group 25-44 is much higher than that of perpetrators in the Bronx. The distribution of the graphs between perpetrator and victim are highly correlated, except for the boroughs of the Bronx and Queens.

<br><br>
The Ratio of Death For The Shooting Incidents In Each Borough (Stacked Bar Chart)
```{r}
ggplot(data, aes(x=BORO, fill=STATISTICAL_MURDER_FLAG)) +
  geom_bar() +
  ggtitle("Shooting incidents in boroughs") +
  labs(x = "Boroughs",y = "Number of Incidence") +
  scale_fill_manual(values = c("#00AFBB", "#FC4E07", "#E7B800"))+
  theme(plot.title = element_text(hjust = 0.5))


```
Bronx and Brooklyn are where most shooting incidents happened, while Bronx has higher death rates than Brooklyn. Manhattan and Queens are similar and relatively safer than the former two Boroughs, having about half less shooting incidents reported. Among all five boroughs, Staten Island is the safest in terms of the total number of shooting incidents.

<br><br>
The Ratio of Death With Respect To The Location of The Shooting Incidents (Cleveland dot plot)
```{r}

df_loc <-  data
df_loc <- df_loc |> filter(STATISTICAL_MURDER_FLAG == "Y" | STATISTICAL_MURDER_FLAG == "N") |> mutate(LOCATION_DESC_new = fct_lump_min(LOCATION_DESC,5))
df_loc <- df_loc |> group_by(LOCATION_DESC_new, STATISTICAL_MURDER_FLAG) |> summarize(n = n())
ggplot(df_loc, aes(x = n, y = fct_reorder2(LOCATION_DESC_new, STATISTICAL_MURDER_FLAG== "Y",n,.desc=FALSE))) + 
  geom_point(aes(color = STATISTICAL_MURDER_FLAG)) +
  #coord_trans(x = "log10") +
  scale_x_continuous(trans = 'log10') +
  labs(
    x = "Number of Incidents",
    y = "Location (Other means incidents < 5)"
  ) +
  scale_color_manual(values = c("#00AFBB", "#FC4E07", "#E7B800"))+
  ggtitle("Shooting Incidents Locations") +
  theme(plot.title = element_text(hjust = 0.5))
  #+ theme(axis.text.x = element_text(angle=30))
```

By plotting the shooting incidents against descriptive locations, we noticed that for most locations the number of incidents resulting in the death of the victim is lower than that of the victim survived, however, for places like beauty/nail salon, social club, bar, and liquor store, this situation is reversed that there are more death cases.

## Visualization Relating to Incidents Date Time

The Ratio of Death in the Shooting Incidents With Respect to Hours In a Day. (Stacked Bar Chart)
```{r}
#df <- data |>
#  pivot_longer(cols = -OCCUR_DATE, names_to = "TYPE", values_to = "RATE") |>
#  mutate(TYPE= forcats::fct_reorder2(TYPE, OCCUR_DATE, RATE))

#df_1 <- data %>% group_by(OCCUR_DATE) %>% mutate(Count = n())
df_1 <- data %>% mutate(OCCUR_TIME = hms(OCCUR_TIME))


#fill="#1380A1"
ggplot(df_1, aes(x=OCCUR_TIME, fill=STATISTICAL_MURDER_FLAG)) +
  geom_histogram(bins=24, color="white") +
  ggtitle("Shooting incidents over time") +
  scale_x_time(breaks = hm(paste0(seq(0, 24, by=4), ":00"))) + 
  labs(x = "",y = "Number of Incidences") +
  scale_fill_manual(values = c("#00AFBB", "#FC4E07", "#E7B800"))+
  theme(plot.title = element_text(hjust = 0.5))
  #+ theme_grey(16)
  #+ theme(legend.title = element_blank())
```
We can observe that shooting incidents have a clear drop between 5:00AM ~ 1:00PM, while the high shooting incident frequency starting around 7:00PM, it reaches peak around 10:00PM. The death cases went up as number of shooting incidents are reported, death rate are not very related to the time of shooting incident happened, however, it is still higher in the evening than daytime. 

<br><br>
The Number of Shooting Incidents With Respect To Month (Line Chart with LOESS Smoother)
```{r}
library(dplyr)
library(plotly)
df_date <- data %>% mutate(OCCUR_DATE = as.Date(OCCUR_DATE, '%m/%d/%Y'))
#str(df_date)

df_date2 <- df_date |> mutate(week = lubridate::week(df_date$OCCUR_DATE)) |> group_by(week) |> summarize(n = n())
df_date2 <- df_date2 |> mutate(Date= lubridate::ymd( "2022-01-01" ) + lubridate::weeks( df_date2$week - 1 ))
#df_date2

#ggplot(df_date, aes(x=format(OCCUR_DATE, "%Y-%m"))) +
ggplot(df_date2, aes(Date, n))+
  geom_line(#color = "cornflowerblue", 
            size=1) +
  geom_point(#color = "cornflowerblue", 
             size=1.3) +
  geom_smooth(method="loess", se=FALSE) +
  labs(x="Date",
       y="Number of Incidents") +
  ggtitle("Shooting Incidents in 2022") +
  theme(plot.title = element_text(hjust = 0.5))
```

From the above graph it is obvious that the number of shooting incidents peaks at July, and there is a general trend that starting from January 2022, the number of incidents went up and reached the peak, then it slightly went down until October, which is the last released time of this data set.

<br><br>
The Relationship Between the Month and Boroughs of Shooting Incidents (Line Chart)
```{r}
library(dplyr)
library(plotly)
df_date <- data %>% mutate(OCCUR_DATE = as.Date(OCCUR_DATE, '%m/%d/%Y'))
#str(df_date)

#df_date2 <-df_date |> mutate(week = cut.Date(OCCUR_DATE, breaks="1 week", labels=FALSE)) |> arrange(OCCUR_DATE)
df_date2 <- df_date |> mutate(week = lubridate::week(df_date$OCCUR_DATE)) |> group_by(week, BORO, .drop = FALSE) |> summarize(n = n())
#df_date2 <- df_date2 |> group_by(OCCUR_DATE, BORO) |> summarize(n = n(), week=week)
df_date2 <- df_date2 |> ungroup() |> mutate(Date= lubridate::ymd( "2022-01-01" ) + lubridate::weeks( df_date2$week - 1 ))
#df_date2

#ggplot(df_date, aes(x=format(OCCUR_DATE, "%Y-%m"))) +
ggplot(df_date2, aes(Date, n, color=BORO))+
  geom_line(aes(group=BORO), size=1) +
  geom_point(size=1.3) +
  labs(x="Date",
       y="Number of Incidents") +
  ggtitle("Shooting Incidents in 2022") +
  theme(plot.title = element_text(hjust = 0.5))

  
#plot_ly(  
#    df_date2, x = ~Date, y = ~n, 
#    type = 'scatter',  
#    mode = 'lines+markers',  
#    # Hover text:  
#    hoverinfo = 'text',  
#    text = ~paste(Date)  
#)  
```
<br>
Alternative (Stacked Histogram)
```{r}
library(dplyr)
library(plotly)
df_date <- data %>% mutate(OCCUR_DATE = as.Date(OCCUR_DATE, '%m/%d/%Y'))
#str(df_date)

#df_date2 <-df_date |> mutate(week = cut.Date(OCCUR_DATE, breaks="1 week", labels=FALSE)) |> arrange(OCCUR_DATE)
df_date2 <- df_date |> mutate(week = lubridate::week(df_date$OCCUR_DATE)) |> group_by(week, BORO, .drop = FALSE)
#df_date2 <- df_date2 |> group_by(OCCUR_DATE, BORO) |> summarize(n = n(), week=week)
df_date2 <- df_date2 |> ungroup() |> mutate(Date= lubridate::ymd( "2022-01-01" ) + lubridate::weeks( df_date2$week - 1 ))
df_date2

#ggplot(df_date, aes(x=format(OCCUR_DATE, "%Y-%m"))) +
ggplot(df_date2, aes(x=Date, fill=BORO))+
  geom_bar()+
  labs(x="Date",
       y="Number of Incidents") +
  ggtitle("Shooting Incidents in 2022") +
  theme(plot.title = element_text(hjust = 0.5))
```
The above two graphs visualized the number of shooting incidents in the five boroughs by weeks in the past 3 quarters. From the line chart we noticed that Brooklyn's incidents peaks at April, but the total number of incidents in all five boroughs peaks at July as shown by the stacked histogram.

##Summary of Results:

In this analysis of the most recent shooting incidents in NYC, we found out that East New York, the Bronx, and Brownsville are the most dangerous areas in New York City in terms of the number of shooting incidents. And black males with age between 25 and 44 are the most vulnerable, as this group of people has the highest number of victims in these incidents. In general, perpetrator tends to be younger than the victims, and the age group 25-44 ranks first in both the number of perpetrators and the number of victims across all boroughs. For college students (aged between 18-24) in NYC, Brooklyn is the most dangerous borough as it has the highest number of victims under this age group. And though in general apartment building has the highest number of shooting incidents, places like Beauty/nail salon, social club, liquor store, and bar have a higher death rate. Speaking of time of the occurence, July is when the number of shooting incidents peak so far this year, while the trend is decreasing since July.

<!--chapter:end:results.Rmd-->


# Interactive component

Placeholder



<!--chapter:end:interactive.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Conclusion

During this process of data analysis, we explored various techniques to visualize the shooting incidents data in New York City and discovered some key findings. We found out that, as expected, there are more male perpetrators/victims than females, and most of the perpetrators or victims are under the age group 25-44. However, we also found out surprisingly that the perpetrators are slightly younger than the victims, meaning that there might be perpetrators targeting older people. As for geography, most shooting incidents occurred in Brooklyn, with the next being the Bronx. And though in general apartment building has the highest number of shooting incidents, places like Beauty/nail salon, social club, liquor store, and bar have a higher death rate. Speaking of time of the occurrence, July is when the number of shooting incidents peak so far this year, while the trend is decreasing since July.

The main limitation of this project is due to the absence of continuous variables in the original data set, even though we tried our best to mutate the data set to get some continuous variables, like the frequency and percentage, however, it still prevent us from using some of the plots for multidimensional continuous variables. Another minor limitation that does not harm the goal of this project but may cause some difficulty to the audience is the large number of rows and columns, which makes the labels on some graphs overlap each other. To overcome this, we adjusted the labels in various ways, like rotating and abbreviating the labels while scaling the axis and hiding the repeating labels, but they still overlap, for example, on the mosaic plot. And abbreviating the labels is not intuitive for interpretation.<br><br>

For future directions, since this is an up-to-date data set that is still updating, it would be helpful to build an API to obtain new shooting incident records and automatically update these visualizations, and publish as a public website for those who may be interested in this topic to view. The data can be stored in a database or put on the cloud, so that every time a shooting incident record is released by the NYPD, the API will be triggered to insert this record into the database.<br><br>

The lesson learned from this exploration is the wide range of possible graphs that we can choose for all types of variables, while most are documented and presented by the professor in class, we also had some research to explore more possible graphs, like the mirror chart and population pyramid, and finding out how some basic graphs, like the bar chart, may be transformed into a visually different graph. <br><br>

<!--chapter:end:conclusion.Rmd-->

